\sektion{5}{When \texorpdfstring{$f^*$}{f*} commutes with finite limits}

 \begin{theorem} \label{lec05T:CommWithFiniteProjLimits}
   If $f:\C'\to \C$ is continuous and finite projective limits are
   representable in $\C'$, then $f^*$ commutes with finite projective limits. In
   particular, $f:T\to T'$ is a morphism of topoi in this case.
 \end{theorem}
 \begin{proof}\anton{this proof needs to be completed and cleaned up}
   By the same argument used to prove Lemma \ref{lec03L:limits_representable}, it is
   enough to show that $f^*$ commutes with products of two sheaves (and thus finite
   products) and equalizers.

%   To say that finite projective limits exist is the same as saying that finite products
%   and fiber products exist \anton{by lemma blah; note that it works for ``finite''
%   everywhere}, which is the same as saying that finite products and equalizers exist. If
%   $F:I\to T$ is a functor and $I$ is finite, you can write $\varprojlim F = Eq\bigl(\prod_{i\in
%   I}F(i)\rightrightarrows \prod_{u\in Mor(I)} F(target(u))\bigr)$. Thus, it is enough to
%   check that $f^*$ commutes with fiber products and products of two sheaves (therefore
%   finite products of sheaves).

   Recall that $f^*F = (\hat f^* F)^a$. One can check that $-^a$ commutes with finite
   projective limits,\footnote{It doesn't commute with infinite products, by the way.} so
   it is enough to show that $\hat f^*:\hat \C\to \hat \C'$ commutes with finite products
   and equalizers.

   (Products) If $F_1,F_2\in \hat \C$, and $U'\in \C'$. Recall the definition of $\hat
   f^*$ from the proof of Proposition \ref{lec04T:left_adjoint}. We have
    \[\xymatrix@R-1.5pc{
     \hat f^*(F_1\times F_2)(U') \ar@{}[d]|{\parallel} & (\hat f^* F_1\times \hat f^*
     F_2)(U') \ar@{}[d]|{\parallel}\\
     \raisebox{-2em}{$\underset{U'\to f(U)}{\varinjlim}F_1(U)\times F_2(U)$} \ar[r]^<>(.5)\Ga &
     \raisebox{-2em}{$\Bigl(\underset{U'\to f(U_1)\hspace{-3ex}}{\varinjlim}F_1(U_1)\Bigr)\times \Bigl(
     \underset{U'\to f(U_2)\hspace{-3ex}}{\varinjlim}F_2(U_2)\Bigr)$} }\] where $\Ga$ is
   defined in the obvious way.\footnote{By the universal properties of $\varinjlim$, it
   is enough to define a map from each $F_1(U)\times F_2(U)$ to some $F_1(U_1)$ and some
   $F_2(U_2)$. Take $U_1=U_2=U$, and then take the obvious projections.} First we check
   injectivity. If $\sigma=\bigl(U'\to f(U),s_1\in F_1(U),s_2\in F_2(U)\bigr)$ and $\tau
   = \bigl(U'\to f(V),t_1\in F_1(V),t_2\in F_2(V)\bigr)$. If $\Ga(\sigma)=\Ga(\tau)$,
   then there is a diagram
   \[\xymatrix{
    U' \ar[r]\ar[d] & f(U_1) \ar[d] \ar[dl]\\
    f(U) & f(V) }\qquad\qquad
    \xymatrix{
    U' \ar[r]\ar[d] & f(U_2) \ar[d] \ar[dl]\\
    f(U) & f(V) }
   \]
   such that $s_1,t_1$ have the same image in $F(U_1)$ and the images of $s_2,t_2$ are
   the same in $F(U_2)$.\anton{if you take a limit over a \emph{filtering} category, then it
   commutes with limits? something is fishy; what is it} We'd like to say that we can
   take $U_1$ and $U_2$ to be the same.

   We have the diagram
   \[\xymatrix{
    U_3 \ar[d]\ar[dr]\\
    U_1\ar[d] & U_2 \ar[dl]\\
    U\times V
   }\]
   That is, we can find some $U_3$ so that blah\anton{}. For surjectivity, you use the
   existence of products.

   (Equalizers) Say we have
   $\xymatrix{F_1\ar@<.5ex>[r]^h \ar@<-.5ex>[r]_g & F_2}$, then we are looking at
   \[\xymatrix @R-1pc{
    \hat f^* Eq\Bigl(F_1 \underset{g}{\overset{h}{\rightrightarrows}} F_2\Bigr)(U')
    \ar@{}[d]|{\parallel}
    &
    Eq\Bigl(
    \hat f^* F_1 \underset{\hat f^*g}{\overset{\hat f^*h}{\rightrightarrows}} \hat f^* F_2
    \Bigr)(U')
    \ar@{}[d]|{\parallel}
   \\
    \raisebox{-2em}{$\underset{U'\to f(U)}{\varinjlim} Eq\Bigl( F_1(U)
   \underset{g(U)}{\overset{h(U)}{\rightrightarrows}} F_2(U)\Bigr)$} \ar[r]
   &
    \raisebox{-2em}{$ Eq\Bigl( \underset{U'\to f(U)}{\varinjlim}F_1(U)
   \underset{\hat f^*g(U')}{\overset{\hat f^*h(U')}{\rightrightarrows}} \underset{U'\to
   f(U)}{\varinjlim}F_2(U)\Bigr)$}
   }\]
    Same sort of arguments to show that this is a bijection.
 \end{proof}
 There are lots of sites that have the property that finite projective limits exist.
 \begin{example}
   $\sch$ has finite projective limits (because we have finite products and fiber
   products). Thus, all big sites have finite projective limits.
 \end{example}
 \begin{example}
   The small \'etale site has finite projective limits. We know how to produce products
   and fiber products
    \[\xymatrix{
    Z\ar@{^(->}[r] \ar[dr] & U_1 \ar@<.5ex>[r] \ar@<-.5ex>[r] \ar[d] & U_2 \ar[dl]\\ & X
   }\]
   The equalizer need not be smooth even if the two maps are smooth, but if the two maps
   are \'etale, then so is the equalizer.
 \end{example}
 \begin{example}
   If $X$ is a scheme, then we get $X_\text{lis-et}\xrightarrow{\varepsilon} X_{et}$ is a
   morphism of topoi, and $\varepsilon_*$ is exact, so you can compute cohomology in
   either topos.
 \end{example}

 \subsektion{Faithfully flat descent}
 If $Y\to X$ is a morphism of schemes, then we get a functor of points
 $h_Y:(\sch/X)^{op}\to \set$ which is a presheaf. The main point is the following.
 \begin{theorem}
   $h_Y$ is a sheaf in the fppf topology (and therefore also in the \'etale topology).
 \end{theorem}
 By the Yoneda embedding, we have $(\sch/X)\hookrightarrow (\sch/X)_{et}=X_{ET}$. We will
 define \emph{algebraic space} to be an object in $X_{ET}$ which is more general than a
 scheme, but in which you can still do geometry (i.e.~where you can redo EGA).

% \begin{tabular}{ll}
%   $X_{zar}$ & small Zariski topos\\
%   $X_{ZAR}$ & big Zariski\\
%   $X_{et}$ & small \'etale topos\\
%   $X_{ET}$ & big \'etale topos\\
%   $X_\text{lis-et}$ & small lisse-\'etale topos\\
%   $X_{fppf}$ & big fppf topos
% \end{tabular}

 \begin{definition}
   A morphism of schemes $f:X\to Y$ is \emph{flat} if for all $x\in X$ the map
   $\O_{Y,f(x)}\to \O_{X,x}$ is flat. It is \emph{faithfully flat} if it is flat and
   surjective. A map of rings $A\to B$ is \emph{faithfully flat} if the map of spectra
   $\spec B\to \spec A$ is.
 \end{definition}
 Key case: $Y=\AA^1_\ZZ$ is flat over $X=\spec \ZZ$.Then
 \[\xymatrix{
 h_Y(\spec A)\ar[r] \ar@{}[d]|{\parallel} &
 h_Y(\spec B)\ar@{}[d]|{\parallel} \ar@<.5ex>[r] \ar@<-.5ex>[r] &
 h_Y(\spec B\otimes_A B) \ar@{}[d]|{\parallel}\\
 A\ar[r] &
 B \ar@<.5ex>[r]^{b\mapsto b\otimes 1} \ar@<-.5ex>[r]_{b\mapsto 1\otimes b} &
 B\otimes_A B
 }\]
 If \anton{iff?} $A\to B$ is faithfully flat, then the bottom sequence is exact.
