 \newpage
 \gdef\sectionname{References}
 \addcontentsline{toc}{section}{References}
 \nocite{Atiyah-Macdonald,
         Conrad,
         EGA,
         FGAex,
         Hartshorne,
         Keel-Mori,
         Knutson,
         Lang:Algebra,
         Raynaud-Gruson,
         SGA,
         Vistoli}

%% These two lines are for producing the bibliography
%  \bibliography{StacksBib}
%  \bibliographystyle{alpha}
%% Then I paste the contents of the bbl file below

 \newcommand{\etalchar}[1]{$^{#1}$}
 \begin{thebibliography}{FGI{\etalchar{+}}05}

\bibitem[AM69]{Atiyah-Macdonald}
M.~F. Atiyah and I.~G. Macdonald.
\newblock {\em Introduction to commutative algebra}.
\newblock Addison-Wesley Publishing Co., Reading, Mass.-London-Don Mills, Ont.,
  1969.

\bibitem[Con]{Conrad}
Brian Conrad.
\newblock The Keel-Mori theorem via stacks.
\newblock Available at \url{www.math.lsa.umich.edu/~bdconrad/}.

 \bibitem[FGI{\etalchar{+}}05]{FGAex}
 Barbara Fantechi, Lothar G{\"o}ttsche, Luc Illusie, Steven~L. Kleiman, Nitin
   Nitsure, and Angelo Vistoli.
 \newblock {\em Fundamental algebraic geometry}, volume 123 of {\em Mathematical
   Surveys and Monographs}.
 \newblock American Mathematical Society, Providence, RI, 2005.
 \newblock Grothendieck's FGA explained.

 \bibitem[EGA]{EGA}
 Alexander Grothendieck and Jean-Alexandre-Eug\`ene Dieudonn\'e.
 \newblock {\em \'{E}l\'ements de g\'eom\'etrie alg\'ebrique.}
 \newblock Numbers 4, 8, 11, 17, 20, 24, 28, and 32. 1960-1967.
 \newblock Institut des Hautes \'Etudes Scientifiques. Publications
   Math\'ematiques.

 \bibitem[SGA]{SGA}
 Alexander Grothendieck, Mich\`ele Raynaud, Michel Demazure, Michael Artin,
   Jean-Louis Verdier, Pierre Deligne, Pierre Berthelot, Luc Illusie, and
   Nicholas Katz.
 \newblock {\em S\'eminaire de G\'eom\'etrie Alg\'ebrique du Bois-Marie}.
 \newblock Springer-Verlag, Berlin, 1970-1973.
 \newblock Lecture Notes in Mathematics, Volumes 151, 152, 153, 224, 225, 269,
   270, 288, 305, 340, 569, and 589.

 \bibitem[Har77]{Hartshorne}
 Robin Hartshorne.
 \newblock {\em Algebraic geometry}.
 \newblock Springer-Verlag, New York, 1977.
 \newblock Graduate Texts in Mathematics, No. 52.

\bibitem[KM97]{Keel-Mori}
Se{\'a}n Keel and Shigefumi Mori.
\newblock Quotients by groupoids.
\newblock {\em Ann. of Math. (2)}, 145(1):193--213, 1997.

 \bibitem[Knu71]{Knutson}
 Donald Knutson.
 \newblock {\em Algebraic spaces}.
 \newblock Springer-Verlag, Berlin, 1971.
 \newblock Lecture Notes in Mathematics, Vol. 203.

 \bibitem[Lan02]{Lang:Algebra}
 Serge Lang.
 \newblock {\em Algebra}, volume 211 of {\em Graduate Texts in Mathematics}.
 \newblock Springer-Verlag, New York, third edition, 2002.

 \bibitem[RG71]{Raynaud-Gruson}
 Michel Raynaud and Laurent Gruson.
 \newblock Crit\`eres de platitude et de projectivit\'e. {T}echniques de
   ``platification'' d'un module.
 \newblock {\em Inventiones Mathematicae}, 13:1--89, 1971.

 \bibitem[Vis05]{Vistoli}
 Angelo Vistoli.
 \newblock Notes on Grothendieck topologies, fibered categories and descent theory.
 \newblock \url{arxiv.org/math.AG/0412512}. This appears as the first part
   of [FGI{\etalchar{+}}05].

 \end{thebibliography}
