\sektion{19}{Sheaf Cohomology} The next results will take a couple of lectures. \begin{theorem}   Let $f:X\to Y$ be a proper morphism of locally noetherian algebraic spaces, and let   $F$ be a coherent sheaf on $X$. Then $R^q\!f_*F$ is a coherent $\O_Y$-module for   all $q\ge 0$. \end{theorem} \begin{remark}   Working man's definition of proper: Let $Y=\spec A$ with $A$ noetherian. Then $X\to   \spec A$ is proper if the following. By Chow's lemma, we can choose $X'\to X$ proper   birational surjection so that $X'\to \spec A$ is projective \anton{proper + birational   implies surjective?}. \anton{something for every choice of $X'$} Then the theorem   roughly says that $H^q(X,F)$ are finitely generated $A$-modules. \end{remark} Something about pullback commuting with finite projective limits wasn't justified. here is a main point. \anton{what?} \begin{lemma}\label{lec18L:preserve_injectives}   Let $F:\A\to \B$ be an additive functor between abelian categories which admits an   exact left adjoint. Then $F$ takes injective objects to injective objects. \end{lemma} \begin{proof}   Let $I\in \A$ be injective, let $0\to M\to N$ be exact in $\A$, and let $G:\B\to \A$   be the left adjoint of $F$. To prove that $F(I)$ is injective, we must show that given   a morphism $M\to F(I)$, we can find the dashed arrow in the diagram.   \[\xymatrix{    0\ar[r] & M\ar[r]\ar[d] & N \ar@{-->}[dl]\\    & F(I)   }\qquad\raisebox{-1pc}{$\xleftrightarrow{\ \text{adjunction}\ }$}\qquad   \xymatrix{    0\ar[r] & G(M) \ar[r]\ar[d] & G(N)\ar@{-->}[dl]\\    & I   }\]    Since $G$ is exact, the top row of the diagram on the right is exact. Since $I$ is   injective, the dashed arrow exists. \end{proof} Let $X$ be an algebraic space, and let $U\to X$ be an \'etale surjection, with $U$ a scheme. For every point $u\in U$, we choose a separable closure of the residue field $k(u)\hookrightarrow \bbar{k(u)}$, and let $\bbar u:(\spec \bbar{k(u)})_{et}\to X_{et}$ be the corresponding point. Since all \'etale covers of a separably closed field are trivial, the topos $(\spec \bbar{k(u)})_{et}$ is the point topos (as a category, it is $\set$), so $\bbar u$ is a point of the topos $X_{et}$. The following lemma roughly says that a sheaf injects into the product of its stalks at these separable points. \begin{lemma}\label{lec18L:enough_points}   Let $X$ be an algebraic space, and let $U\to X$ be an \'etale cover by a scheme. For   any sheaf $F\in X_{et}$, one gets an injective map of sheaves $F\to \prod_{u\in U}   \bbar u_* \bbar u^* F$. \end{lemma} \begin{proof}   Let $(V\to X)\in Et(X)$ and $\alpha_1,\alpha_2 \in F(V)$ which map the same element   in the product $\prod_u \bbar u_* \bbar u^* F$. Then we want to show that they are   equal.   It is enough to show that if $\alpha_1,\alpha_2\in F(X)$, then stuff \dots, that is,   we may assume $V=X$ for some reason. Moreover, we can assume $X$ is a   scheme.\anton{how do we make these reductions?}   {$\bbar u_* \bbar u^* F(V) = \bbar u^* F(V\times_X \spec \bbar{k(u)})$. This   is the sheaf associated to the presheaf where we compute $\varinjlim_{W} F(W)$,   where the limit is over $W$ such that   \[\xymatrix{   V\times_X \spec \bbar{k(u)} \ar@{}[r]|<>(.5){=} & \spec \bbar{k(u)} \ar[r] \ar[dr] & W\ar[d]\\   & & X   }\]}   So for every $u\in U$, there exists an \'etale morphism $W_u\to X$ whose image   contains the image of $u$ such that $\alpha_1|_{W_u}=\alpha_2|_{W_u}$. That means   that $\coprod W_u\to X$ is an \'etale surjection and $\alpha_1=\alpha_2$ when   restricted to the cover $\coprod W_u$. That means that they are equal because $F$ is   a sheaf. \end{proof} \begin{lemma}\label{lec18L:enough_injectives_among_modules}   Let $X$ be an algebraic space, and let $\Lambda$ be a sheaf of rings in $X_{et}$. Then   the category $\Lambda\mod$ has enough injectives.   %of $\Lambda$-modules $Mod_\Lambda(X_{et})$ \end{lemma} This is really true much more generally by the way. There will be enough injectives whenever there are ``enough points'' in the sense of Lemma \ref{lec18L:enough_points}. \begin{proof}   If $F$ is a $\Lambda$-module, then $\bbar u^* F$ is a $\bbar u^* \Lambda$-module. We   know that the category of modules over a ring has enough injectives, so choose an   embedding $\bbar u^* F\hookrightarrow I_{\bbar u}$ where $I_{\bbar u}$ is an injective   $\bbar u^* \Lambda$-module. Then we have $F\hookrightarrow \prod_{\bbar u} \bbar   u_*\bbar u^* F\hookrightarrow \prod_{\bbar u} \bbar u_* I_{\bbar u}$. Built into the   definition of a morphism of topoi is that $u^*$ is exact, so Lemma   \ref{lec18L:preserve_injectives} tells us that $\bbar u_* I_{\bbar u}$ is injective,   and an arbitrary product of injective objects is injective. \end{proof} The key point to the next theorem will be that if $X$ is an affine scheme, and $F$ is a quasi-coherent sheaf in $X_{et}$, then $H^q(X_{et},F)=0$ for $q>0$. This is good to prove if you're bored with the lectures.